---
title: "P8105 HW2 nk3037"
author: "Navya Koneripalli"
date: "2023-10-01"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

# Question 1
### Cleaning pols-month.csv
```{r, message = FALSE}
month_df = 
  tibble(
    month_num = 1:12,
    month = month.name
  )

pols_month_data = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -prez_dem, -prez_gop)
```

### Cleaning snp.csv
```{r, message=FALSE}
snp_data = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
  arrange(year, month) %>% 
  mutate(month = month.name[month])
```

### Cleaning unemployment.csv
```{r, message=FALSE}
unemployment_data =
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  rename(year = Year) %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
  )
```

### Merging pols_month_data, snp_data and unemployment_data
```{r, message=FALSE}
merged_data =
  pols_month_data %>% 
  left_join(snp_data) %>% 
  left_join(unemployment_data)
```

The `merged_data` dataset contains `r nrow(merged_data)` rows and `r ncol(merged_data)` columns.

The `pols_month_data` has data on the political party association of the president, governors and senators for a given year. The dataset contains `r nrow(pols_month_data)` rows and `r ncol(pols_month_data)` columns. It contains data from the years `r range(pols_month_data$year)[1]` to `r range(pols_month_data$year)[2]`.

The `snp_data` dataset contains data related to the S&P stock market index. It contains `r nrow(snp_data)` rows and `r ncol(snp_data)` columns.It has data for the year `r range(snp_data$year)[1]` to `r range(snp_data$year)[2]`.

Lastly, the `unemployment_data` dataset contains data on the monthly unemployment rate. It has `r nrow(unemployment_data)` rows and `r ncol(unemployment_data)` columns for the years `r range(unemployment_data$year)[1]` to `r range(unemployment_data$year)[2]`.