---
title: "P8105 HW2 nk3037"
author: "Navya Koneripalli"
date: "2023-10-01"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

# Question 1
### Cleaning pols-month.csv
```{r, message = FALSE}
month_df = 
  tibble(
    month_num = 1:12,
    month = month.name
  )

pols_month_data = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -prez_dem, -prez_gop)
```

### Cleaning snp.csv
```{r, message=FALSE}
snp_data = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
  arrange(year, month) %>% 
  mutate(month = month.name[month])
```

### Cleaning unemployment.csv
```{r, message=FALSE}
unemployment_data =
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  rename(year = Year) %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
  )
```

### Merging pols_month_data, snp_data and unemployment_data
```{r, message=FALSE}
merged_data =
  pols_month_data %>% 
  left_join(snp_data) %>% 
  left_join(unemployment_data)
```

The `merged_data` dataset contains `r nrow(merged_data)` rows and `r ncol(merged_data)` columns.

The `pols_month_data` has data on the political party association of the president, governors and senators for a given year. The dataset contains `r nrow(pols_month_data)` rows and `r ncol(pols_month_data)` columns. It contains data from the years `r range(pols_month_data$year)[1]` to `r range(pols_month_data$year)[2]`.

The `snp_data` dataset contains data related to the S&P stock market index. It contains `r nrow(snp_data)` rows and `r ncol(snp_data)` columns.It has data for the year `r range(snp_data$year)[1]` to `r range(snp_data$year)[2]`.

Lastly, the `unemployment_data` dataset contains data on the monthly unemployment rate. It has `r nrow(unemployment_data)` rows and `r ncol(unemployment_data)` columns for the years `r range(unemployment_data$year)[1]` to `r range(unemployment_data$year)[2]`.

# Question 2
### Data manipulation for Mr. Trash Wheel dataset
```{r}
mr_trash_wheel = 
  read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N586")
mr_trash_wheel = janitor::clean_names(mr_trash_wheel, case = "snake") %>% #convert all variable names into snake_case

mutate(homes_powered = (weight_tons * (500/30)))
# Calculated the average number of homes powered for each tonne of trash and applied the formula to all the rows of the Mr. Trash Wheel dataset

mr_trash_wheel$year = as.numeric(mr_trash_wheel$year)
# Converting the character year variable into a numeric one so that it can be binded with the year variables from the other datasets

mr_trash_wheel = mr_trash_wheel%>% 
  mutate(trash_wheel_name = "Mr. Trash Wheel")
# Added an additional variable trash_wheel_name to help with the final bind of the datasets

mr_trash_wheel = mr_trash_wheel %>% 
  select(-sports_balls, -glass_bottles)
# removed columns so that the number and column headings match with the other two datasets for ease of binding
```

### Data manipulation for Professor Trash Wheel dataset
```{r}
prof_trash_wheel =
  read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M108")
prof_trash_wheel = janitor::clean_names(prof_trash_wheel, case = "snake") %>% #convert all variable names into snake_case

mutate(homes_powered = (weight_tons * (500/30)))
# Calculated the average number of homes powered for each tonne of trash and applied the formula to all the rows of the Professor Trash Wheel dataset

prof_trash_wheel = prof_trash_wheel%>% 
  mutate(trash_wheel_name = "Professor Trash Wheel")
# Added an additional variable trash_wheel_name to help with the final bind of the datasets

prof_trash_wheel = prof_trash_wheel %>% 
  select(-glass_bottles)
# removed column so that the number and column headings match with the other two datasets for ease of binding
```

### Data manipulation for Gwynnda Trash Wheel dataset
```{r}
gwynnda_trash_wheel =
  read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", range = "A2:L157")
gwynnda_trash_wheel = janitor::clean_names(gwynnda_trash_wheel, case = "snake") %>% #convert all variable names into snake_case

mutate(homes_powered = (weight_tons * (500/30)))
# Calculated the average number of homes powered for each tonne of trash and applied the formula to all the rows of the Gwynnda Trash Wheel dataset

gwynnda_trash_wheel = gwynnda_trash_wheel%>% 
  mutate(trash_wheel_name = "Gwynnda Trash Wheel")
# Added an additional variable trash_wheel_name to help with the final bind of the datasets
```

### Merging Mr. Trash Wheel, Prof. Trash Wheel and Gwynnda Trash Wheel datasets
```{r}
merged_trash_wheel_part1 = rbind(mr_trash_wheel, prof_trash_wheel)
merged_trash_wheel = rbind(merged_trash_wheel_part1, gwynnda_trash_wheel)
# binded the three datasets on top of each other

filtered_mr_trash = merged_trash_wheel %>% 
  filter(trash_wheel_name == "Mr. Trash Wheel")
# filtered the dataset to contain only Mr. Trash Wheel to calculate the sum of weight of trash collected.

filtered_gwynnda = merged_trash_wheel %>% 
  filter(trash_wheel_name == "Gwynnda Trash Wheel" & month == "July" & year == "2021")
# filtered the dataset to contain only Gwynnda Trash Wheel in July 2021 to calculate the sum of cigarettes collected.
```

The `merged_trash_wheel` dataset contains `r nrow(merged_trash_wheel)` rows and `r ncol(merged_trash_wheel)` columns. 

The merged dataset contains trash collection data for the years between `r range(merged_trash_wheel$year)[1]` and `r range(merged_trash_wheel$year)[2]`. 

The `merged_trash_wheel`dataset captures the weight (in tons), volume and count of each individual type of waste collected by each of the 3 trash wheels by dumpster number. 

The average weight of trash collected by all of the trash wheels across the full time period of the dataset is `r mean(merged_trash_wheel$weight_tons)` tons. 

The total weight of trash collected by Mr. Trash Wheel is `r sum(filtered_mr_trash$weight_tons)` tons.

The total number of cigarettes collected by Gwynnda Trash Wheel in July 2021 is `r sum(filtered_mr_trash$cigarette_butts)` cigarette butts.

# Question 3
### Data manipulation of mci_baseline data
```{r, message = FALSE, warning=FALSE}
mci_baseline =
  read_csv("./data/data_mci/MCI_baseline.csv", skip = 1) #reading the .csv file into R
mci_baseline = janitor::clean_names(mci_baseline, case = "snake") # convert all variable names into snake_case

mci_baseline = mci_baseline %>% 
  mutate(
    sex = recode(sex, "1" = "Male", "0" = "Female"), 
    apoe4_status = recode(apoe4, "1" = "Carrier", "0" = "Non-carrier")
    ) #recoded the sex and apoe4_status to non-numeric variable types

mci_baseline$age_at_onset = as.numeric(mci_baseline$age_at_onset) # converted the age_at_onset variable from character to numeric so that the dataset can be filtered in the next step

mci_baseline_excluded = mci_baseline %>% 
  filter((age_at_onset - current_age) > 0)
# removing the patients that do not meet the requirements
```

The `mci_baseline` dataset contains the data of `r nrow(mci_baseline)` patients. Of this, `r nrow(mci_baseline_excluded)` patients had developed MCI. The average age of the patients was `r mean(mci_baseline$current_age)` years. The dataset contained `r nrow (filter(mci_baseline, sex == "Female"))` females. Among the women, `r nrow(filter(mci_baseline, sex == "Female" & apoe4_status == "Carrier"))` were carriers. Of the females, `r nrow(filter(mci_baseline_excluded, sex == "Female"))` developed MCI.

### Data manipulation of mci_amyloid data
```{r, message = FALSE, warning = FALSE}
mci_amyloid = 
  read.csv("./data/data_mci/mci_amyloid.csv", skip = 1) #reading the .csv file into R
mci_amyloid = janitor::clean_names(mci_amyloid, case = "snake") # convert all variable names into snake_case

mci_amyloid = mci_amyloid %>%
  rename(baseline_post_2hrs= time_2) %>% 
  rename (baseline_post_4hrs= time_4) %>% 
  rename(baseline_post_6hrs= time_6) %>% 
  rename(baseline_post_8hrs= time_8)
#renamed the columns to be more descriptive of the variable

mci_amyloid = mci_amyloid %>% 
  rename(id = study_id)
#renamed the id variable so that it is consistent with mci_baseline

columns_to_convert <- c("baseline", "baseline_post_2hrs", "baseline_post_4hrs", "baseline_post_6hrs", "baseline_post_8hrs")

mci_amyloid[columns_to_convert] <- lapply(mci_amyloid[columns_to_convert], as.numeric)
# convert the variable types into numeric
```

The `mci_amyloid` dataset contains the data of `r nrow(mci_amyloid)` patients and `r ncol(mci_amyloid)`columns. Of this, `r nrow(mci_amyloid)` patients had MCI at baseline. 

The average baseline amyloid ratio is `r mean(mci_amyloid$baseline, na.rm = TRUE)`. 

The average amyloid ratio 2 hours post baseline is `r mean(mci_amyloid$baseline_post_2hrs, na.rm = TRUE)`. 

The average amyloid ratio 4 hours post baseline is `r mean(mci_amyloid$baseline_post_4hrs, na.rm = TRUE)`.

The average amyloid ratio 6 hours post baseline is `r mean(mci_amyloid$baseline_post_6hrs, na.rm = TRUE)`. 

The average amyloid ratio 8 hours post baseline is `r mean(mci_amyloid$baseline_post_8hrs, na.rm = TRUE)`. 